Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-17 cordillera-cycle.tex
Reformat and improve notes and todos.

% My commands
\newcommand{\idea}[1]{\textbf{[IDEA: #1]}}
\newcommand{\note}[1]{\textbf{[NOTE: #1]}}
\newcommand{\todo}[1]{\textbf{[TODO: #1]}}
\newcommand{\aref}[0]{\textbf{[ref.]}}

@@

\begin{center}
\begin{tabular}{>{\bfseries}ll}
    \tophline
    Introduction & Happy with it except for missing refs.\\
    Model setup  & Happy with it except for missing refs.\\
    Results      & Happy with it except for missing refs.\\
    Discussion   & Deglaciation part needs improvement.\\
    Conclusions  & Can be compacted and the style improved.\\
    \bottomhline
\end{tabular}
\end{center}

@@

For more than a century, exploration and geological investigation of the
Cordillera have led to a large collection of evidence of the former ice cover
\citep{Jackson.Clague.1991}. This evidence consists of mapped boundaries of the
former ice extent (\citealp[Fig.~1.12]{Clague.1989}; \citealp{Duk-Rodkin.1999};
\citealp{Dyke.2004}), directions of past sliding velocities
(\citealp{Prest.etal.1968}; \citealp[Fig.~2]{Kleman.etal.2010}), and locations
of former
melt-water streams \citep{Margold.etal.2011, Margold.etal.2013}, as well as the
timing of glaciation identified from radiocarbon dating \citep[e.g.,][]
{Clague.1981, Porter.Swanson.1998}, cosmogenic dating \citep[e.g.,][]
{Ward.etal.2007, Menounos.etal.2009, Stroeven.etal.2010, Stroeven.etal.2013},
and the offshore sedimentary record \citep{Cosma.etal.2008, Davies.etal.2011}.

\note{Is the set of references in the above paragraph representative of the
      current state of the field? Am I missing something important here?}

@@

A~pseudo-plastic sliding law \citep{PISM-authors.2014} \note{ask Ed Bueler if
this will be published somewhere soon} relates the
bed-parallel shear stress and the sliding velocity. The yield stress is
modelled using the Mohr--Coulomb criterion. The till friction angle $\phi$
varies from 15 to 45{\degree}. It is taken as a~function of modern bed
elevation, with lowest values occurring at low elevation, thus accounting
for a~weakening of till associated with the presence of marine sediments.
The till effective pressure is related to the modelled amount of water in the
till, using a formula derived from laboratory experiments with till extracted
from an Antarctic ice stream \citep[Eqn.~2]{Tulaczyk.etal.2000}. Basal
topography is
derived from the ETOPO1 combined topography and bathymetry dataset with
a~resolution of 1\,arc-min \citep{Amante.Eakins.2009}. Sea level is lowered as
a function of time based on the Spectral Mapping Project (SPECMAP)
\citep[SPECMAP,][]{Imbrie.etal.1989} time scale.
Basal topography responds to ice load
following a bedrock deformation model that includes point-wise isostasy,
elastic lithosphere flexure and viscous mantle deformation in a~semi-infinite
half-space \citep{Lingle.Clark.1985,Bueler.etal.2007}.

@@

Independently of the choice of palaeo-temperature record used as forcing
time-series for the ice sheet model, our simulations consistently produce two
glaciation maxima during MIS~4 (61.9--55.4\,kyr) and MIS~2
(29.5--16.9\,kyr). These events broadly correspond in timing to the
Gladstone (MIS~4) and McConell (MIS~2) glaciations documented by geological
evidence in the northern part of the model domain \aref, and to the Fraser
(MIS~2) glaciation in the southern part \aref.

\note{Could you help me to fill the missing refs in this last sentence?}

@@

However, our simulations depict, independently of the choice of record used for
palaeo-temperature forcing, a central Skeena mountain ice cap persisting
throughout the entire last glacial cycle (Fig.~\ref{fig:snapshots} and
\ref{fig:duration}). In particular, this ice cap persists during MIS~3,
providing a nucleation centre for the onset of glacial
readvance towards the last glacial maximum (MIS~2).

@@

In addition, the erosional landscape of the Skeena mountain bears one of the
strongest glacial imprint (?) found within the area formerly covered by the
Cordilleran ice sheet \aref. Here we suggest that persistent ice cover
associated with basal ice temperatures at the pressure-melting point
(Fig.~\ref{fig:warmbase} and ~\ref{fig:warmbase}) may partly explain this
spectacular landscape.

\note{Can we use a reference here to support the point of strong glacial
      imprint in the Skeena Mountains?}

@@

% fig:warmbase
\begin{figure*}
  \includegraphics{warmbase}
  \caption{Modelled fraction of warm-based ice cover during the ice-covered
           period. Note the dominance of warm-based conditions on the
           continental shelf and major glacial troughs of the coastal ranges.
           Hatches indicate areas that were covered by cold ice only.}
  \label{fig:warmbase}
\end{figure*}

% fig:warmfrac
\begin{figure*}
  \includegraphics{warmfrac}
  \caption{Modelled duration of warm-based ice cover during the last
           120\,kyr. Long ice cover durations combined with basal
           temperatures at the pressure-melting point may explain the strong
           glacial erosional imprint of the Skeena Mountains (SM) landscape.
           Hatches indicate areas that were covered by cold ice only.
           \note{indicate location of Skeena Mountains.}}
  \label{fig:warmfrac}
\end{figure*}

% fig:erosion
\begin{figure*}
  \includegraphics{erosion}
  \caption{Modelled cumulative basal displacement (integrand of basal velocity)
           over the last 120\,kyr.
           \note{I am still unsure about the use of this figure. I may
                 remove it from this paper and keep the idea for later use.}}
  \label{fig:erosion}
\end{figure*}

>>>>

% My commands
\newcommand{\note}[1]{\textbf{[NOTE: #1]}}
\newcommand{\todo}[1]{\emph{[\textbf{Todo:} #1]}}
\newcommand{\aref}[0]{\textbf{[ref.]}}

@@

@@

For more than a century, exploration and geological investigation of the
Cordillera have led to a large collection of evidence of the former ice cover
\citep{Jackson.Clague.1991}. This evidence consists of mapped boundaries of the
former ice extent (\citealp[Fig.~1.12]{Clague.1989}; \citealp{Duk-Rodkin.1999};
\citealp{Dyke.2004}), directions of past sliding velocities
(\citealp{Prest.etal.1968}; \citealp[Fig.~2]{Kleman.etal.2010}), and locations
of former
melt-water streams \citep{Margold.etal.2011, Margold.etal.2013}, as well as the
timing of glaciation identified from radiocarbon dating \citep[e.g.,][]
{Clague.1981, Porter.Swanson.1998}, cosmogenic dating \citep[e.g.,][]
{Ward.etal.2007, Menounos.etal.2009, Stroeven.etal.2010, Stroeven.etal.2013},
and the offshore sedimentary record \citep{Cosma.etal.2008, Davies.etal.2011}%
\footnote{\textbf{Arjen, Martin,}, please comment on my choice of references
    in this paragraph. Is it a fair sample of the current state of the field?
    Am I missing some other important works?}.

@@

A~pseudo-plastic sliding law \citep{PISM-authors.2014}\footnote{I need to ask
    Ed Bueler if the sliding will be published somewhere soon.} relates the
bed-parallel shear stress and the sliding velocity. The yield stress is
modelled using the Mohr--Coulomb criterion. The till friction angle $\phi$
varies from 15 to 45{\degree}. It is taken as a~function of modern bed
elevation, with lowest values occurring at low elevation, thus accounting
for a~weakening of till associated with the presence of marine sediments.
The till effective pressure is related to the modelled amount of water in the
till, using a formula derived from laboratory experiments with till extracted
from an Antarctic ice stream \citep[Eqn.~2]{Tulaczyk.etal.2000}. Basal
topography is
derived from the ETOPO1 combined topography and bathymetry dataset with
a~resolution of 1\,arc-min \citep{Amante.Eakins.2009}. Sea level is lowered as
a function of time based on the Spectral Mapping Project (SPECMAP)
\citep[SPECMAP,][]{Imbrie.etal.1989} time scale.
Basal topography responds to ice load
following a bedrock deformation model that includes point-wise isostasy,
elastic lithosphere flexure and viscous mantle deformation in a~semi-infinite
half-space \citep{Lingle.Clark.1985,Bueler.etal.2007}.

@@

Independently of the choice of palaeo-temperature record used as forcing
time-series for the ice sheet model, our simulations consistently produce two
glaciation maxima during MIS~4 (61.9--55.4\,kyr) and MIS~2
(29.5--16.9\,kyr). These events broadly correspond in timing to the
Gladstone (MIS~4) and McConell (MIS~2) glaciations documented by geological
evidence in the northern part of the model domain \aref, and to the Fraser
(MIS~2) glaciation in the southern part \aref
\footnote{\textbf{Arjen, Martin, Johan,} can you check that I use these names
    correctly, and help me to fill the missing refs in this sentence? Is is
    possible to trace back when these names where first applied?}.

@@

However, our simulations depict, independently of the choice of record used for
palaeo-temperature forcing, a central Skeena mountain ice cap persisting
throughout the entire last glacial cycle (Figs.~\ref{fig:snapshots}
and~\ref{fig:duration}). In particular, this ice cap persists during MIS~3,
providing a nucleation centre for the onset of glacial
readvance towards the last glacial maximum (MIS~2).

@@

In addition, the erosional landscape of the Skeena mountain bears one of the
strongest glacial imprint found within the area formerly covered by the
Cordilleran ice sheet \aref%
\footnote{\textbf{Arjen, Martin, Johan,} can we use a reference here to support
    this point (observations, mapping, shaded relief)? Alternatively, we could
    maybe use a photograph, if any of you has one. It would add a bit of
    variation into the figure list. \textbf{Johan}, you seemed to support this
    this point during our discussion, any idea?}.
Here we suggest that persistent ice cover (Fig.~\ref{fig:duration})
associated with basal ice temperatures at the pressure-melting point
(Fig.~\ref{fig:warmfrac} and ~\ref{fig:warmbase}%
\footnote{Fig.~\ref{fig:erosion} is tricky because it may be significantly
    affected by the constant-geothermal heat flux assumption, which is not
    valid for this region. Arjen, Johan and I decided to leave it for my kappa
    for now. Regarding Figs.~\ref{fig:warmfrac} and ~\ref{fig:warmbase}, I
    still wonder if it is useful to show both. Basically we have
    Fig.~\ref{fig:warmbase}\,=\,Fig.~\ref{fig:duration}\,*\,Fig.~\ref{fig:warmfrac}.
    Maybe the reader can perform this multiplication visually without need of
    Fig.~\ref{fig:warmbase}? Tell me what you think of it.})
may partly explain this spectacular landscape.

@@

% fig:warmfrac
\begin{figure*}
  \includegraphics{warmfrac}
  \caption{Modelled fraction of warm-based ice cover during the ice-covered
           period. Note the dominance of warm-based conditions on the
           continental shelf and major glacial troughs of the coastal ranges.
           Hatches indicate areas that were covered by cold ice only.
           \todo{indicate location of the Skeena Mountains.}}
  \label{fig:warmfrac}
\end{figure*}

% fig:warmbase
\begin{figure*}
  \includegraphics{warmbase}
  \caption{Modelled duration of warm-based ice cover during the last
           120\,kyr. Long ice cover durations combined with basal
           temperatures at the pressure-melting point may explain the strong
           glacial erosional imprint of the Skeena Mountains (SM) landscape.
           Hatches indicate areas that were covered by cold ice only.
           \todo{indicate location of the Skeena Mountains.}}
  \label{fig:warmbase}
\end{figure*}

% fig:erosion
\begin{figure*}
  \includegraphics{erosion}
  \caption{Modelled cumulative basal displacement (integrand of basal velocity)
           over the last 120\,kyr.
           \todo{remove this figure (see footnote in text).}}
  \label{fig:erosion}
\end{figure*}
